---
title: "SQL_assigment_HadasKaiser"
format: html
editor: visual
---

## Setting up DB

```{r}
library(ggplot2)
library(dplyr)
library(dbplyr)

con_chocolate <- dbConnect(RSQLite::SQLite(), dbname = "chocolate.sqlite")

tables <- dbListTables(con_chocolate)
print(tables)

# Extract each table as a df
customers_df <- dbGetQuery(con_chocolate, "SELECT * FROM customers")
orders_df <- dbGetQuery(con_chocolate, "SELECT * FROM orders")
products_df <- dbGetQuery(con_chocolate, "SELECT * FROM products")
salesreps_df <- dbGetQuery(con_chocolate, "SELECT * FROM salesreps")
```

## Q1

1.  Assuming you got it right - what did your boss want? Explain the query.
2.  Run the query - what is the answer?

```{r}
Q1 <- dbGetQuery(con_chocolate, "
SELECT salesreps.Name, SUM(orders.amount) AS total_candy_bars_sold
FROM salesreps INNER JOIN orders USING (srid)
WHERE
orders.sale_date BETWEEN '2022-01-01' AND '2022-12-31'
AND salesreps.year_joined = 2010
GROUP BY salesreps.Name
")
```

**Explaining the query:** The following query identifies the number of candy bars sold in 2022 by sales representatives who joined the company in 2010.

The query's results is the following table:

| Name                | Total candy bar sold |
|---------------------|----------------------|
| Tootle, Naudia      | 1825                 |
| al-Farrah, Ghaaliba | 1866                 |
| al-Sadri, Saamyya   | 1784                 |

## Q2

Generate a frequency table of the number of times a customer made an order. It should have two columns: 1. total_orders - the number of orders made by a customer.\
2. N - the number of customers that made total_orders orders.

```{r}
Q2 <- dbGetQuery(con_chocolate, "
    select total_orders, count(cid) as N
    from
    (
    SELECT cid, count(cid) as total_orders
    FROM orders
    GROUP BY cid
    order by count(cid)
    )
    group by 1
")
```

## Q3

In a single query:

• Identify the candy bar with the highest number of total orders in the first quarter of 2022.

• What is the name of the sales rep that sold the most of those candy bars in the second quarter of 2022?

```{r}
Q3 <- dbGetQuery(con_chocolate, "
with FirstQ AS (
    select pid,sum(amount) as total_orders
    FROM orders
    where sale_date between '2022-01-01' and '2022-03-31'
    group by 1
    order by 2 desc
    limit 1
),
SecondQ AS (
    select srid, pid, sum (amount) as total_sold
    from orders
    where sale_date between '2022-04-01' and '2022-06-30'
    group by 1,2
),
TopSales AS (
    select sq.srid, sq.total_sold
    from SecondQ as sq
    left join FirstQ as fq on fq.pid = sq.pid
    order by 2 desc
    limit 1
)

select sr.name
from TopSales ts
left join salesreps sr
on sr.srid = ts.srid
")
```
